\begin{lec}[2011-12-08]\end{lec}
\stepcounter{section}
\tocsubsection{Minimum Cost Flows}

\begin{defn}
Let $D = (V, A)$ be a digraph with arc capacities $c(a) \geq 0$ for all $a
\in A$ and cost coefficients $w(a)$ for all $a \in A$. Let $s, t \in V$.

Consider all $(s, t)$-flows of value $f$. The \new{Minimum Cost Flow (MCF)
Problem}{Minimaler Kosten Netzwerkflussproblem} consists of finding a
$(s,t)$-flow $x$ with $val(x) = f$ and cost $\sum_{a \in A} w(a) x_a$
minimum among all $(s,t)$-flows of value $f$.
\end{defn}

The MCF Problem can be formulated as linear program (later more) and special
network-simplex algorithms exist to solve the problem in polynomial time.
Alternatively several combinatorial algorithms exist.

\begin{defn}
Let $x$ be a feasible $(s,t)$-flow in $D$ and let $C$ be a (not necessarily
directed) cycle in $D$. The cycle $C$ can be orientated in two possible ways
(clockwise or counterclockwise). Given an orientation of $C$, \emph{forward
arcs} on $C$ are directed along the orientation, \emph{backwards arcs} opposite.

A cycle $C$ is called an \new{augmenting cycle}{augmentierender Kreis} w.r.t. $x$ if there exists an
orientation of $X$ such that
\begin{align*}
x_a < c_a & \text{ for all forward arcs } a \in C \\
x_a > 0	& \text{ for all backward arcs } a \in C
\end{align*}
\end{defn}

\begin{defn}
A \emph{cost} of an augmenting cycle $C$ is defined as
\[
\sum_{a \in C: forward} w_a - \sum_{a \in c: backward} w_a
\]
\end{defn}

\begin{thm}\label{thm:thm164}
A feasible $(s,t)$-flow $x$ in $D$ with value $f$ has minimum cost if and
only if no augmenting cycle $C$ w.r.t. $x$ with negative costs exists.
\end{thm}
\begin{proof}
only $\Rightarrow$, $\Leftarrow$ later.

Let $\sum w(a) x_a$ be minimum. Assume that there exists an augmenting cycle
with negative cost. We define
\[
	\varepsilon = \min ( c_{ij} - x_{ij} if (i,j) \in C forward ; x_{ij} if
	(i,j) \in C backward )
\]
and 
\[
	\tilde{x}_{ij} = ( x_{ij} + \varepsilon if (i,j) \in C forward, 
	x_{ij} - \varepsilon if (i,j) \in C backward,
	x_{ij} if (i,j) \in A \setminus C
	)
\]
Now, $\tilde{x}$ is a feasible $(s,t)$-flow with $val(\tilde{x}) = val(x) =
f$ and cost
\[
	\sum_{a \in A} w(a) \tilde{x}_a = \sum_{a \in A} w(a) x_a + \varepsilon
	\left( \sum_{a \in C: foward} w(a) - \sum_{a \in C: backward} w(a) \right)
	< \sum_{a \in A} w(a) x_a
\]

To proe the backward direction, we first describe the algorithmic procedure.
\end{proof}

\begin{defn}
Given a $(s,t)$-flow $x$ with $val(x) = f$, we define the \new{augmenting
network}{augmentierendes Netzwerk} w.r.t. $x$ as follows: $N = (V, \overline{A}, \overline{C},
\overline{w})$ with
\begin{align*}
\overline{A} &= A_1 \cup A_2 \\
A_1 &= \{ ij \in A: x_{ij} < c_{ij} \} \\
A_2 &= \{ ji: ij \in A, x_{ij} > 0 \}
\end{align*}

For $a \in A$, we denote by $a_1 \in A_1$ and / or $a_2 \in A_2$ the
corresponding arcs in $A_1$ and $A_2$. If $a_1$ (or $a_2$) does not exist,
we evaluate $x(a_1)$ (or $x(a_2)$) with zero.
Further,
\begin{align*}
\overline{c}(\overline{a}) = &
	\begin{cases}
		c(a) - x(a) & \text{if } \overline{a} = a_1 \\
		x(a) & \text{if } \overline{a} = a_2
	\end{cases} & \forall a \in A \\
\overline{w}(\overline{a}) = &
	\begin{cases}
		 w(a) & \text{if } \overline{a} = a_1 \\
		-w(a) & \text{if }\overline{a} = a_2
	\end{cases}   & \forall a \in A
\end{align*}
\end{defn}

\begin{lem}
Every augmenting cycle in $D$ corresponds with exactly one \emph{directed}
cycle in $N$. The cost of both cycles are identical.
\end{lem}
\begin{proof}
Exercise sheet.
\end{proof}

\begin{thm}
The flow $x$ is cost minimal among all $(s,t)$-flows in $D$ of value $f$ if
and only if no directed cycle in $N$ has negative cost.
\end{thm}
\begin{proof}
$\Rightarrow$: analogue to Theorem~\ref{thm:thm164}.

$\Leftarrow$: Assume $x$ is not cost minimal. Hence, there exists a
$\tilde{x}$ with value $f$ and $w^T\tilde{x} < w^Tx$.
Now define for $\overline{a} \in \overline{A}$ (w.r.t. $x$)
\[
	\overline{x}(\overline{a}) = ( max \{ 0, \tilde{x}(a) - x(a) \} if
	\overline{a} = a_1 \in A_1, max \{0, x(a) - \tilde{x}(a)\} if
	\overline{a} = a_2 \in A_2 )
\]
Thus it holds that $\overline{x}(a_1) - \overline{x}(a_2) = \tilde{x}(a) -
x(a)$. Further, it holds that 
\[
	\overline{x}(a_1) \overline{w}(a_1) + \overline{x}(a_2) \overline{w}(a_2)
	= w(a) ( \tilde{x}(a) - x(a) )
\]
and thus
\[
	\sum_{\overline{a} \in \overline{A}} \overline{w}(\overline{a})
	\overline{x}(\overline{a}) = 
	\sum_{a \in A} w(a) ( \tilde{x}(a) - x(a) ) = w^T \tilde{x} - w^Tx < 0
\]

Moreover, flow conservation holds for all $v \in V$:
\begin{align*}
	&\sum_{a \in \delta^+_N(v)} \overline{x}(a) - \sum_{a \in \delta^-_N(v)}
	\overline{x}(a) \\
	= &\sum_{a \in \delta^+_D(v)} \overline{x}(a_1) -
	\overline{x}(a_2) - \sum_{a \in \delta^-_D(v)} \overline{x}(a_1) -
	\overline{x}(a_2) \\
	= &\sum_{a \in \delta^+(v)} \tilde{x}(a) - x(a) - \sum_{a \in
	\delta^-(v)} \tilde{x}(a) - x(a) \\
	= &\left( \sum_{a \in \delta^+(v)} \tilde{x}(a) - \sum_{a \in
	\delta^-(v)} \tilde{x}(a) \right) - 
	\left( \sum_{a \in \delta^+(v)} x(a) - \sum_{a \in
	\delta^-(v)} x(a) \right) \\
	= &( 0 - 0 if v \in V \setminus \{s,t\},
	f - f = 0 if v = s,
	-f + f = 0 if v = t )
\end{align*}
Thus $\overline{x}$ is a feasible flow of value $0$ and
$\overline{w}^T\overline{x} < 0$.
Since $\overline{x} \neq 0$ and flow conservation holds for all $v \in V$,
$\overline{x}$ is a \new{circulation}{Zirkulation} and can be decomposed
into flows $x_i$ defined on cycles $C_i$ ($i = 1, ..., k < |\overline{A}|$).
With $\sum_{i=1}^k x_i(\overline{a}) = \overline{x}(\overline{a}) \forall
\overline{a} \in \overline{A}$.

For at least one cycle $C_j$ it must hold that $\overline{w}^T x_j < 0$,
otherwise $\overline{w}^T \overline{x} \geq 0$. Then the vector $x_j$
defines an augmenting cycle with negative cost.
Contradiction.
\end{proof}
